Enabling Ambient Backscatter \\Using Low-Cost Software Defined Radio

### 2017-04-07 paper.tex
Worked on evaluation and discussion.

\section{Evaluation}
In this section we present our evaluation results. The \textit{RTL-SDR}, combined with intelligent signal processing in \textit{Octave}, are the main tools we have employed.

@@

We have used the textit{RTL-SDR} to observe the space variations of the signal strength. \textit{Octave} scripts have been written to aid in this purpose. These scripts are available online under \url{https://github.com/s3xm3x/RTLSDRspecAn}. The scripts perform a frequency sweep through the desired band. The obtained samples are then transferred from the time domain to the frequency domain using an FFT.

Because the RTL-SDR has a maximum sampling rate of 1.8 MSPs, we are
limited to a maximum simultaneous acquisition band of 900 kHz. Because
of this, we sample the desired frequency band at 900 kHz intervals and
stitch these bands together to form the overall frequency sweep. This
approach is valid only because we are not interested in real-time data.
With this approach we are able to scan a 20 MHz band in roughly two
minutes.\cp{Please confirm these things.}

During the scanning process	the gain is set to 40.2 dB. To get the
signal strength we simply take the average over a 10 MHz band around the
center frequency of the desired TV signal. This approach can be justified by the fact, that DVB-T is specified up to 10 MHz bandwidth. And the local TV signal provider is to its own statements using DVB-T 2. The signal \ensuremath{|H|} is calculated as follows.
\begin{equation}
	|H| = \Biggl| \sum_{n=0}^N FFT\bigl( \Re\{ s_{\text{band,n}}(t) \} \bigr) \Biggr|
\end{equation}     
Where \ensuremath{N} is the total number of FFT bins accumulated over
the entire frequency sweep\cp{Please confirm I got this right}. The FFT
is only carried out over the real values, which are received from the
\textit{RTL-SDR}. \ensuremath{s_{\text{band,n}}(t)} is the signal in the
time domain of one band. There are as many samples taken in one band as
needed for an FFT of size 2048. No window is used when stitching
individual scans together. One can see, the effects of this a little bit, when taking a closer look at the plotted data. Rising and falling edges at the beginning of each band are visible. To get \ensuremath{|H_{tv,avg}|} one has to normalize everything and calculate the power. Before doing that a simple smoothing algorithm (FIR) is applied on \ensuremath{|H|}.
\begin{equation}
	|H_{tv,avg}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
\end{equation}   
where \ensuremath{|H_0|} is a normalization value measured at a certain
point in space. 
\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/haversine}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
	\end{minipage}
	\caption{Signal strength fading against distance relative to the spot where the maximum signal strength has been captured (TV towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and 594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been calculated with the Haversine formula.}
\label{fig:haversine}
\end{figure}

\subsection{Signal Strength Variation over Time}

\subsection{Communication Performance}
We tested the communication with different data rates from 1 bit/s up to 1 kbit/s. As one can see in figure \ref{fig:transmission} the signal strength is really low after backscattering. This leads to the problem, that the signal has a really bad quality. This related to the fact, that not all bits of the A/D converter are used. Hence a lot of quantization noise is introduced. In other words more gain is needed. The transmission shown in figure \ref{fig:transmission} was carried out with the maximum gain of 50 dB, which is available with the in \ref{sub:rtl2832} decribed tuner. So we got to the limits of the \textit{RTL-SDR} at that point. We were able to achieve a range indoors with 1 bit/s of a couple of meteres, but when we use a higher bitrate we can only communicate over a range of a couple of decimeter and still have a bit error ratio of approximately 40 \%. The communiction experiment was also carried out outside closer to the tower. We could not figure out any improvements going closer to the tower.  
\section{Discussion}
We are aware of the fact, that our communication system has to be improved until it is able to be used in allday technology. These are aspects, where fine tuning is necessary to enhance the performance:
\begin{itemize}
	\item Antennas: Different backscattering antennas have to be tried out to find the one with the best results. The question is mainly to which frequency the antenna has to be tuned (e.g. to the center of the TV signal or more to the edge of it). Moreover the standard \textit{RTL-SDR} antenna including the coaxial cable are of quite low quality. The antenna is e.g. to short. 
	\item Receiver: Some fine tuning can be made when it comes to the filter coefficients of the different filters.
	\item Frame: The frame structure can be optimized to be able to correct errors (e.g. Hamming code).  
\end{itemize}   
To come to a conclusion one can say, that we were able
\begin{itemize}
	\item to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
	\item to provide a heat map of tv signals in a mid-sized swedish city as well as the tools to create it.
	\item to show for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
\end{itemize}

@@

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

>>>>

\section{Evaluation}
In this section we present our evaluation results. The \textit{RTL-SDR},
combined with intelligent signal processing in \textit{Octave}, are the
main tools we have employed for measurements.

@@

We have used the textit{RTL-SDR} to observe the space variations of the
signal strength. \textit{Octave} scripts have been written to aid in
this purpose. These scripts are available online under
\url{https://github.com/s3xm3x/RTLSDRspecAn}. The scripts perform a
frequency sweep through the desired band. The obtained samples are then
transferred from the time domain to the frequency domain using a fast
fourier transformation (FFT).

Because the RTL-SDR has a maximum (stable) sampling rate of 2.4 MSPs, we
are limited to a maximum simultaneous acquisition band of 1.2 MHz. We
are keep a safety margin to the maximum capabilities and sample the
desired frequency band at 900 kHz intervals and stitch these bands
together to form the overall frequency sweep. This approach is valid
only because we are not interested in real-time data.  With this
approach we are able to scan a 20 MHz band in roughly two minutes.

During the scanning process the gain is set to 40.2 dB. To get the
signal strength we simply take the average over a 10 MHz band around the
center frequency of the desired TV signal. This approach can be
justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
And the local TV signal provider is to its own statements using DVB-T 2.
The signal \ensuremath{|H|} is calculated as follows.  
\begin{equation}
	|H| = \Biggl| \sum_{n=0}^N FFT\bigl( \Re\{ s_{\text{band,n}}(t) \} \bigr) \Biggr|
\end{equation}     
where \ensuremath{N} is the total number of intervals accumulated over
the entire frequency sweep. The FFT
is only carried out over the real values, which are received from the
\textit{RTL-SDR}. \ensuremath{s_{\text{band,n}}(t)} is the signal in the
time domain of one interval (one subband). There are as many samples taken in one band as
needed for an FFT of size 2048. To finally get \ensuremath{|H_{tv,avg}|} one has to normalize everything and calculate the power. Before doing that a simple smoothing algorithm (FIR) is applied on \ensuremath{|H|}.
\begin{equation}
	|H_{tv,avg}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
\end{equation}   
where \ensuremath{|H_0|} is a normalization value measured at a certain
point in space. 
\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/haversine}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
	\end{minipage}
	\caption{Left: Signal strength fading against distance relative
to the spot where the maximum signal strength has been captured (TV
towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been
calculated with the Haversine formula. Right: Heatmap of the 626 MHz TV
signal in Uppsala.}

\label{fig:haversine}
\end{figure}

\subsection{Communication Performance}
We tested the communication with different data rates from 1 bit/s up to
1 kbit/s. As one can see in figure \ref{fig:transmission} the signal
strength is quite low after backscattering. So the problem appears, that
not all bits of the A/D converter are used. Hence a lot of quantization
noise is introduced. In other words more gain is needed. The
transmission shown in figure \ref{fig:transmission} was carried out with
the maximum gain of 50 dB, which is available with the in
\ref{sub:rtl2832} decribed tuner. So we got to the limits of the
\textit{RTL-SDR} at that point. We were able to achieve a range indoors
with 1 bit/s of a couple of meteres, but when we use a higher bitrate we
can only communicate over a range of a couple of decimeter and still
have a bit error ratio of approximately 40 \%. The communiction
experiment was also carried out outside closer to the tower. We could
not figure out any improvements going closer to the tower.  
\section{Discussion}
We are aware of the fact, that our communication system has to be improved until it is able to be used in allday technology. These are aspects, where fine tuning is necessary to enhance the performance:
\begin{itemize}
	\item Antennas: Different backscattering antennas have to be
tried out to find the one with the best results. The question is mainly
to which frequency the antenna has to be tuned (e.g. to the center of
the TV signal or more to the edge of it). Moreover the standard
\textit{RTL-SDR} antenna including the coaxial cable are of quite low
quality. The antenna is e.g. too short. 
	\item Receiver: Some fine tuning can be made when it comes to the filter coefficients of the different filters.
	\item Frame: The frame structure can be optimized to be able to correct errors (e.g. Hamming code).  
\end{itemize}   
To come to a conclusion one can say, that we were able
\begin{itemize}
	\item to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
	\item to provide a heat map of TV signals in a mid-sized swedish city as well as the tools to create it.
	\item to show for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
\end{itemize}

@@

% use section* for acknowledgment
%\section*{Acknowledgment}

%The authors would like to thank...
